The Birth of a Galaxy: Primordial Metal Enrichment and Stellar Populations

### 2011-08-26 ApJ/ms.tex
Minor corrections after a proofreading.

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{ABN02, Bromm02_P3, OShea07a}.  A fraction of these
stars enrich the surrounding intergalactic medium (IGM) when they go
supernova, which can happen in stars $\lsim 40~\Ms$ in Type II
supernovae (SNe) or in stars roughly between 140 \Ms~and 260 \Ms~in
pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.  The host
halo and the neighboring halos are then enriched with this ejecta.
There exists a critical metallicity that is $\sim 10^{-6} Z_\odot$ if
dust cooling is efficient \citep{Omukai05, Schneider06_Frag, clark08}
and $\sim 10^{-3.5} Z_\odot$ otherwise \citep{Bromm01,
  2009ApJ...691..441S}, where the gas can cool rapidly, lowering its
Jeans mass.  An intermediate characteristic mass of $\sim10~\Ms$ can
be occur if the gas cooling is suppressed to the cosmic microwave
background (CMB) temperature \citep{Larson98, Tumlinson07_IMF,
  2009ApJ...691..441S}.  The resulting Population II star cluster will
thus have a lower characteristic stellar mass than its metal-free
progenitors.  These first stellar clusters may be connected to stars
in the Milky Way halo and nearby dwarf spheroidal (dSph) galaxies,
both with a metallicity floor of [Z/H]\footnote{We use the
  conventional notation, [Z/H] $\equiv \log(Z/H) -
  \log(Z_\odot/H_\odot)$.} = --4 \citep{Beers05, Tafelmeyer10,
  Frebel10_Obs}.

@@

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  Spurred by the notion of a
critical metallicity, this transition has been extensively studied
with (i) volume-averaged semi-analytic models \citep{Scannapieco03,
  Yoshida04, Furlanetto05_Reion}, (ii) models using hierarchical
merger trees \citep{Tumlinson06, Tumlinson10, Salvadori07, Komiya10},
(iii) post-processing of cosmological simulations with blastwave
models \citep{Karlsson08, Trenti09, Trenti10}, and (iv) direct
numerical simulations with stellar feedback \citep{Tornatore07,
  Ricotti08, Maio10_Pop32, Maio10_Enrich}.

@@

We use the adaptive mesh refinement (AMR) code
enzo~v2.0\footnote{\texttt{enzo.googlecode.com, changeset
    b86d8ba026d6}} \citep{OShea2004}.  It uses an $N$-body adaptive
particle-mesh solver \citep{Efstathiou85} to follow the dark matter
(DM) dynamics.  It solves the hydrodynamical equation using the
second-order accurate piecewise parabolic method \citep{Woodward84,
  Bryan95}, while a Riemann solver ensures accurate shock capturing
with minimal viscosity.  We use the recently added HLLC Riemann solver
\citep{Toro94_HLLC} for additional stability in strong shocks and
rarefaction waves.  We use the nine-species (\ion{H}{1}, \ion{H}{2},
\ion{He}{1}, \ion{He}{2}, \ion{He}{3}, e$^-$, \hh, \hh$^+$, H$^-$)
non-equilibrium chemistry model in \enzo~\citep{Abel97, Anninos97} and
the \hh~cooling rates from \citet{Glover08_Rates}.  We spatially
distinguish metal enrichment from Pop II and Pop III stars.  We will
follow-up this study with one that considers radiative cooling from
metal-enriched gas, using the method of \citet{2008MNRAS.385.1443S}
that is already implemented in \enzo.

To resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior \citep[also
  see][]{OShea08}.  We also refine on a DM overdensity of $3 \times
2^l$ and always resolve the local Jeans length by at least 4 cells,
avoiding artificial fragmentation during gaseous collapses
\citep{Truelove97}.  If any of these critera are met in a single cell,
it is flagged for further spatial refinement.

@@

For Pop III stars, we use the same SF model as \citet{Wise08_Gal}
where each star particle represents a single star, forming at an
overdensity of $5 \times 10^5$ ($\sim 10^3\; \cubecm$ at $z=10$).  If
multiple cells meet the star particle formation criteria within 1 pc,
we form one Pop III star particle with the center of mass of these
flagged cells to ensure that massive star is created per metal-free
molecular cloud.  Instead of using a fixed stellar mass like in our
previous work, we randomly sampled from an IMF with a functional form
of
%
\begin{equation}
\label{eqn:imf}
f(\log M) dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm
      char}}{M}\right)^{1.6}\right] dM
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the number
of times the random number generator (Mersenne twister;
\citet{MTwister}) has been called for use when restarting the
simulations.

@@

We treat Pop II star formation with the same prescription as
\citet{Wise09}, which is similar to the Pop III prescription but
without the \hh~fraction floor that is removed because the
metal-enriched gas can efficiently cool even in the presence of a
strong UV radiation field \citep[e.g.][]{Safranek10}.  This method is
effectively a modified version of the \citet{Cen92} method but
accounts for the fact that the molecular clouds are resolved.  In the
simulation, we define the molecular cloud as a sphere, centered on the
most dense cell, with a dynamical time $t_{\rm dyn} = 3$ Myr
(corresponding to an average density $\bar{\rho}_{\rm MC} \simeq
1000\mu \cubecm$) and a radius $R_{\rm MC}$, where $\mu$ is the mean
molecular weight.  In this sphere, a fraction $c_\star = 0.07 f_{\rm
  cold}$ of the cold gas ($T < 10^3$ K) is converted into a star
particle with mass $m_\star = c_\star (4\pi/3) \bar{\rho}_{\rm MC}
R_{\rm MC}^3$.  We then replace the sphere with a uniform density
$\rho_{\rm MC} = (1 - c_\star) / (Gt_{\rm dyn}^2)$ and temperature $T
= 10^4$ K, which approximates the initial stages of an \ion{H}{2}
region.  Similar to the Pop III treatment, we merge any newly created
particles within a radius $R_{\rm merge}$.  We set the minimum mass of
a star particle to $m_{\rm \star, min} = 1000 \Ms$.  If the initial
mass does not exceed $m_{\rm \star, min}$, the star particle does not
provide any feedback and continues to accrete until it reaches $m_{\rm
  \star, min}$.  The star particles live for 20 Myr, the maximum
lifetime of an OB star.  These stars generate the majority of the
ionizing radiation and SNe feedback in stellar clusters, thus we
ignore any feedback from lower mass stars.

@@

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime and $E_{\rm ph} = 21.6$ eV, appropriate for a [Z/H] =
$-1.3$ population.  We note that lower-metallicity stars generate up
to 60\% more ionizing photons and might be underestimating the
radiative feedback \citep{Schaerer03}.  Nonetheless this study
provides an excellent first insight in the transition to Pop II SF, as
the metal enrichment is the key ingredient here.  For SN feedback,
these stars generate $6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$
after living for 4 Myr, which is injected into spheres with a radius
of 10 pc, using the same technique as the Pop III SNe.  However if the
resolution of the grid is less than 10/3 pc, we distribute the energy
into a $3^3$ cube surrounding the star particle.  In the same region,
star particles also return ejected material with mass
%
\begin{equation}
  \label{eqn:SNmass}
  \Delta m_{\rm ej} = \frac{0.25 \; \Delta t \; M_\star} 
         {t_0 - 4\;\mathrm{Myr}}
\end{equation}
back to the grid at every timestep on the finest AMR level.  Here
$t_0$ = 20 Myr is the lifetime of the star particle.  This ejected gas
has solar metallicity $Z = 0.02$, resulting in a total metal yield $y
= 0.005$.  In addition to numerical stability, energy distribution
across several cells instead of a single cell has been shown to match
cosmic SFRs better in galaxy simulations at lower redshifts
\citep{Smith10}.

@@

The cosmic Pop II SFR is best fit with an exponential $\exp(-z/1.1)$.
Metal-enriched star formation begins early in the simulation at $z
\sim 16$ when a PISN blastwave overtakes a satellite halo with $M \sim
5 \times 10^5 \Ms$, triggering star formation there.  At $z>10$, these
enriched minihalos cause the Pop II SFR to be higher than this
relation.  Only after $z \sim 10$ does the SFR follows this
exponential, which is a steeper time dependency than the $\exp(-z/3)$
relation found in \citet{Hernquist03}.  Note that they also
acknowledged that this behavior should occur at early times.  At
redshift 7, it reaches $2 \times 10^{-2}$ \sfr.  This value agrees
with derived SFRs from $z=6-10$ galaxy luminosity functions
\citep[e.g.][]{Bouwens10} and large-scale galaxy formation simulations
\citep[e.g.][]{Schaye10}.

@@

Mass-to-light ratios are crucial in interpreting observations when
connecting the galaxy to its DM halo, and ultimately to models of
galaxy formation.  For this paper, we use a simple model to calculate
the total stellar luminosity of each galaxy.  In a later paper, we
will follow up with more sophisicated models to determine more
accurate observables.  Stars in each particle are treated an
instantaneous burst.  Using the empirical law of main sequence
lifetime $t_{\rm MS} \propto M^{-5/2}$, the maximum stellar mass in
such a system is dependent only on its age $t_\star$, so $M_{\rm up} =
(t_\star / 10\;\mathrm{Gyr})^{-2/5}$.  Then combining another scaling
law $L \propto M^{7/2}$ and the Salpeter IMF $\psi(M) \propto
M^{-\alpha}$,
%
\begin{eqnarray}
  \label{eqn:lum}
  L/M_\odot &=& \int_{M_{\rm lo}}^{M_{\rm up}} \psi(M) L(M) dM
  \nonumber\\ &\simeq& M_{\rm up}^{9/2-\alpha} (9/2 - \alpha)^{-1},
\end{eqnarray}
where we set $M_{\rm lo} \equiv 0$ because it is negligible in the
luminosity.  Then the galaxy luminosity is the sum of all star
particle luminosities, which we plot in Fig. \ref{fig:massfn} with
respect to halo mass, along with the DM halo mass function.

>>>>

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{ABN02, Bromm02_P3, OShea07a}.  A fraction of these
stars enrich the surrounding intergalactic medium (IGM) when they go
supernova, which can happen in stars $\lsim 40~\Ms$ in Type II
supernovae (SNe) or in stars roughly between 140 \Ms~and 260 \Ms~in
pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.  The host
halo and the neighboring halos are then enriched with this ejecta.
There exists a critical metallicity that is $\sim 10^{-6} Z_\odot$ if
dust cooling is efficient \citep{Omukai05, Schneider06_Frag, clark08}
and $\sim 10^{-3.5} Z_\odot$ otherwise \citep{Bromm01,
  2009ApJ...691..441S}, where the gas can cool rapidly, lowering its
Jeans mass.  An intermediate characteristic mass of $\sim10~\Ms$ can
be occur if the gas cooling is suppressed to the cosmic microwave
background (CMB) temperature \citep{Larson98, Tumlinson07_IMF,
  2009ApJ...691..441S}.  The resulting Population II star cluster will
thus have a lower characteristic stellar mass than its metal-free
progenitors.  These first stellar clusters may be connected to stars
in the Milky Way halo and nearby dwarf spheroidal (dSph) galaxies,
both with a metallicity floor of [Z/H]\footnote{We use the
  conventional notation, [Z/H] $\equiv$ log(Z/H) -
  log(Z$_\odot$/H$_\odot$).} = --4 \citep{Beers05, Tafelmeyer10,
  Frebel10_Obs}.

@@

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion, halo
mergers, and nearby stellar feedback.  Spurred by the notion of a
critical metallicity, this transition has been extensively studied
with (i) volume-averaged semi-analytic models \citep{Scannapieco03,
  Yoshida04, Furlanetto05_Reion}, (ii) models using hierarchical
merger trees \citep{Tumlinson06, Tumlinson10, Salvadori07, Komiya10},
(iii) post-processing of cosmological simulations with blastwave
models \citep{Karlsson08, Trenti09, Trenti10}, and (iv) direct
numerical simulations with stellar feedback \citep{Tornatore07,
  Ricotti08, Maio10_Pop32, Maio10_Enrich}.

@@

We use the adaptive mesh refinement (AMR) code
\enzo~v2.0\footnote{\texttt{enzo.googlecode.com, changeset
    b86d8ba026d6}} \citep{OShea2004}.  It uses an $N$-body adaptive
particle-mesh solver \citep{Efstathiou85} to follow the dark matter
(DM) dynamics.  It solves the hydrodynamical equation using the
second-order accurate piecewise parabolic method \citep{Woodward84,
  Bryan95}, while a Riemann solver ensures accurate shock capturing
with minimal viscosity.  We use the recently added HLLC Riemann solver
\citep{Toro94_HLLC} for additional stability in strong shocks and
rarefaction waves.  We use the nine-species (\ion{H}{1}, \ion{H}{2},
\ion{He}{1}, \ion{He}{2}, \ion{He}{3}, e$^-$, \hh, \hh$^+$, H$^-$)
non-equilibrium chemistry model in \enzo~\citep{Abel97, Anninos97} and
the \hh~cooling rates from \citet{Glover08_Rates}.  We spatially
distinguish metal enrichment from Pop II and Pop III stars.  We will
follow-up this study with one that considers radiative cooling from
metal-enriched gas, using the method of \citet{2008MNRAS.385.1443S}
that is already implemented in \enzo.

To resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior \citep[also
  see][]{OShea08}.  We also refine on a DM overdensity of three and
always resolve the local Jeans length by at least 4 cells, avoiding
artificial fragmentation during gaseous collapses \citep{Truelove97}.
If any of these critera are met in a single cell, it is flagged for
further spatial refinement.

@@

For Pop III stars, we use the same SF model as \citet{Wise08_Gal}
where each star particle represents a single star, forming at an
overdensity of $5 \times 10^5$ ($\sim 10^3\; \cubecm$ at $z=10$).  If
multiple cells meet the star particle formation criteria within 1 pc,
we form one Pop III star particle with the center of mass of these
flagged cells to ensure that massive star is created per metal-free
molecular cloud.  Instead of using a fixed stellar mass like in our
previous work, we randomly sample from an IMF with a functional form
of
%
\begin{equation}
\label{eqn:imf}
f(\log M) dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm
      char}}{M}\right)^{1.6}\right] dM
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the number
of times the random number generator (Mersenne twister;
\citet{MTwister}) has been called for use when restarting the
simulations.

@@

We treat Pop II star formation with the same prescription as
\citet{Wise09}, which is similar to the Pop III prescription but
without the \hh~fraction floor that is removed because the
metal-enriched gas can efficiently cool even in the presence of a
strong UV radiation field \citep[e.g.][]{Safranek10}.  This method is
effectively a modified version of the \citet{Cen92} method but
accounts for the fact that the molecular clouds are resolved.  In the
simulation, we define the molecular cloud as a sphere, centered on the
most dense cell, with a dynamical time $t_{\rm dyn} = 3$ Myr
(corresponding to an average density $\bar{\rho}_{\rm MC} \simeq
1000\mu \cubecm$) and a radius $R_{\rm MC}$, where $\mu$ is the mean
molecular weight.  In this sphere, a fraction $c_\star = 0.07 f_{\rm
  cold}$ of the cold gas ($T < 10^3$ K) is converted into a star
particle with mass $m_\star = c_\star (4\pi/3) \bar{\rho}_{\rm MC}
R_{\rm MC}^3$.  We then replace the sphere with a uniform density
$\rho_{\rm MC} = (1 - c_\star) / (Gt_{\rm dyn}^2)$ and temperature $T
= 10^4$ K, which approximates the initial stages of an \ion{H}{2}
region.  Similar to the Pop III treatment, we merge any newly created
particles within a radius of 10 pc.  We set the minimum mass of a star
particle to $m_{\rm \star, min} = 1000 \Ms$.  If the initial mass does
not exceed $m_{\rm \star, min}$, the star particle does not provide
any feedback and continues to accrete until it reaches $m_{\rm \star,
  min}$.  The star particles live for 20 Myr, the maximum lifetime of
an OB star.  These stars generate the majority of the ionizing
radiation and SNe feedback in stellar clusters, thus we ignore any
feedback from lower mass stars.

@@

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime and $E_{\rm ph} = 21.6$ eV, appropriate for a [Z/H] =
$-1.3$ population.  We note that lower-metallicity stars generate up
to 60\% more ionizing photons and might be underestimating the
radiative feedback \citep{Schaerer03}.  Nonetheless this study
provides an excellent first insight in the transition to Pop II SF, as
the metal enrichment is the key ingredient here.  For SN feedback,
these stars generate $6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$
after living for 4 Myr, which is injected into spheres with a radius
of 10 pc.  However if the resolution of the grid is less than 10/3 pc,
we distribute the energy into a $3^3$ cube surrounding the star
particle.  In the same region, star particles also return ejected
material with mass
%
\begin{equation}
  \label{eqn:SNmass}
  \Delta m_{\rm ej} = \frac{0.25 \; \Delta t \; M_\star} 
         {t_0 - 4\;\mathrm{Myr}}
\end{equation}
back to the grid at every timestep on the finest AMR level.  Here
$t_0$ = 20 Myr is the lifetime of the star particle.  This ejected gas
has solar metallicity $Z = 0.02$, resulting in a total metal yield $y
= 0.005$.  In addition to numerical stability, energy distribution
across several cells instead of a single cell has been shown to match
cosmic SFRs better in galaxy simulations at lower redshifts
\citep{Smith10}.

@@

The cosmic Pop II SFR is best fit with an exponential $\exp(-z/1.1)$.
Metal-enriched star formation begins early in the simulation at $z
\sim 16$ when a PISN blastwave overtakes a satellite halo with $M \sim
5 \times 10^5 \Ms$, triggering star formation there.  At $z>10$, these
enriched minihalos cause the Pop II SFR to be higher than this
relation.  Only after $z \sim 10$ does the SFR follows this
exponential, which is a steeper time dependency than the $\exp(-z/3)$
relation found in \citet{Hernquist03}, who also acknowledged that this
behavior should occur at early times.  At redshift 7, it reaches $2
\times 10^{-2}$ \sfr.  This value agrees with derived SFRs from
$z=6-10$ galaxy luminosity functions \citep[e.g.][]{Bouwens10} and
large-scale galaxy formation simulations \citep[e.g.][]{Schaye10}.

@@

Mass-to-light ratios are crucial in interpreting observations when
connecting the galaxy to its DM halo, and ultimately to models of
galaxy formation.  For this paper, we use a simple model to calculate
the total stellar luminosity of each galaxy.  In a later paper, we
will follow up with more sophisicated models to determine more
accurate observables.  Stars in each particle are treated an
instantaneous burst.  Using the empirical law of main sequence
lifetime $t_{\rm MS} \propto M^{-5/2}$, the maximum stellar mass in
such a system is dependent only on its age $t_\star$, so $M_{\rm up} =
(t_\star / 10\;\mathrm{Gyr})^{-2/5}$.  Then combining another scaling
law $L \propto M^{7/2}$ and the Salpeter IMF $\psi(M) \propto
M^{-\alpha}$,
%
\begin{eqnarray}
  \label{eqn:lum}
  L/M_\odot &=& \int_{M_{\rm lo}}^{M_{\rm up}} \psi(M) L(M) dM
  \nonumber\\ &\simeq& M_{\rm up}^{9/2-\alpha} (9/2 - \alpha)^{-1},
\end{eqnarray}
where we set $M_{\rm lo} \equiv 0$ because the total luminosity of
low-mass stars is negligible.  Then the galaxy luminosity is the sum
of all star particle luminosities, which we plot in
Fig. \ref{fig:massfn} with respect to halo mass, along with the DM
halo mass function.
